How fluid-mechanical erosion creates anisotropic porous media

### 2022-12-15 docs/PhysicsPaper/Referee Reports/Ref 1 Response/Ref1.tex
NM

Of course, in our manuscript, we consider the Stokes limit of the drag, for which there is a different formula, but the principle remains that it can be decomposed into a dimensional component (that depends on the length-scale of the object, fluid properties, the flow speed, etc.) and a dimensionless coefficient to represent the detailed shape of the object.
Likewise, the permeability associated with a single object can be represented in a similar way, where the dimensionless coefficient could be considered the `shape' permeability, $k_{11}^{(s)}$, analogous to the shape anisotropy defined in our manuscript. That is, the value $k_{11}^{(s)}$ associated with a single body depends on the detailed shape of that body, and not on its overall size, or the viscosity of the fluid, or the flow speed, etc. The dimensional component of the permeability we can call `configurational' permeability $k_{11}^{(c)}$, which, though it seems strange for a single body, will connect to the configurational anisotropy. The value $k_{11}^{(c)}$ represents the dimensional part of the permeability, and it takes into account the length-scale of the body, the strength of the flow, the viscosity of the fluid etc. Therefore, in this notation, we have
\begin{equation}
k_{11} = k_{11}^{(s)} k_{11}^{(c)}
\end{equation}
These quantities could of course be defined similarly for a collection of objects, as in our paper. It is important to note that, in the extraction of $k_{11}^{(c)}$ in our paper, we replace every body with a circle of the same location {\em and the same area}. The later point implies that $k_{11}^{(c)}$ represents the component of $k_{11}$ that depends on overall body length-scale, so that what is left, $k_{11}^{(s)}$ does {\em not} depend on length scale, and instead only depends on the detailed shapes of the bodies.

Introducing a similar decomposition of transverse permeability $k_{22}$ and putting everything together yields
\begin{equation}
\anis = \frac{k_{11}}{k_{22}} = \frac{ k_{11}^{(s)} k_{11}^{(c)} }{ k_{22}^{(s)} k_{22}^{(c)} }
= \anis^{(s)} \anis^{(c)}
\end{equation}
which is the result used in our paper. We have added a brief explanation in the updated manuscript at the end of Section 4.3.
\\ \\

@@

Great question. This is an aspect of the simulation that we had regarded as too detailed to fully explain in this manuscript, but it is an interesting aspect that we are happy to explain here. In essence, we use a simulated annealing procedure to initialize the configuration of circular bodies. We first determine the distribution of circle radii by straightforward sampling (we use a Chi distribution, but this choice is arbitrary) and by using the porosity that is specified. Second, we randomly assign the circle centers; many of the circles will overlap (unless we are incredibly lucky). Next, we specify fictitious forces that act to repel overlapping bodies and also repel bodies from the domain walls. In addition, the force acting on each body has a stochastic component with magnitude, or `temperature', that decreases over the course of the initialization routine. Hence, early in the initialization routine, the temperature is high and the random component is large compared to the deterministic forces, but at later stages, the temperature is low and the random component tends to zero. This algorithm allows the bodies to jump around substantially to explore different configurations and eventually converge to a configuration in which none overlap. Naturally, there are limitations to such an initialization algorithm. For example, there exists thresholds for the solid area fraction (that depend on the radius distribution) above which no circle packing is possible.
We have added a brief explanation of this initialization algorithm to the updated manuscript at the end of Section 3.2.

>>>>

Of course, in our manuscript, we consider the Stokes limit of the drag, for which there is a different formula, but the principle remains that it can be decomposed into a dimensional component (that depends on the length-scale of the object, fluid properties, the flow speed, etc.) and a dimensionless coefficient to represent the detailed shape of the object.
Likewise, the permeability associated with a single object can be represented in a similar way, where the dimensionless coefficient, e.g.~the $C_D$, could be considered the `shape' permeability, $k_{11}^{(s)}$. That is, the value $k_{11}^{(s)}$ associated with a single body depends on the detailed shape of that body, and not on its overall size, or the viscosity of the fluid, or the flow speed, etc. The dimensional component of the permeability we can call `configurational' permeability $k_{11}^{(c)}$, which, though it seems strange for a single body, will connect to the configurational anisotropy. The value $k_{11}^{(c)}$ represents the dimensional part of the permeability, and it takes into account the length-scale of the body, the strength of the flow, the viscosity of the fluid etc. Therefore, in this notation, we have
\begin{equation}
k_{11} = k_{11}^{(s)} k_{11}^{(c)}
\end{equation}
These quantities could of course be defined similarly for a collection of objects, as in our paper. It is important to note that, in the extraction of $k_{11}^{(c)}$ in our paper, we replace every body with a circle of the same location {\em and the same area}. The later point implies that $k_{11}^{(c)}$ represents the component of $k_{11}$ that depends on overall body length-scale, so that what is left, $k_{11}^{(s)}$ does {\em not} depend on length scale, and instead only depends on the detailed shapes of the bodies.

Introducing a similar decomposition of transverse permeability $k_{22}$ and putting everything together yields
\begin{equation}
\anis = \frac{k_{11}}{k_{22}} = \frac{ k_{11}^{(s)} k_{11}^{(c)} }{ k_{22}^{(s)} k_{22}^{(c)} }
= \anis^{(s)} \anis^{(c)}
\end{equation}
which is the result used in our paper. We have added a brief explanation of all this in the updated manuscript at the end of Section 4.3.
\\ \\

@@

Great question. This is an aspect of the simulation that we had initially regarded as too detailed to fully explain in the manuscript, but it is an interesting aspect that we are happy to explain here, and we have added a brief explanation in the updated manuscript. In essence, we use a simulated annealing algorithm to initialize the configuration of circular bodies. We first determine the distribution of circle radii by straightforward sampling (we use a Chi distribution, but this choice is arbitrary). Second, we randomly assign the circle centers; many of the circles will overlap, unless we are incredibly lucky. Next, we specify fictitious forces that act to repel overlapping bodies and also repel bodies from the domain walls. In addition, we imbue this force with a stochastic component with magnitude, or `temperature', that decreases over the course of the initialization routine. Hence, early in the initialization routine, the temperature is high and the random component is large compared to the deterministic forces, but at later stages, the temperature is low and the random component tends to zero. This algorithm allows the bodies to jump around substantially to explore different configurations and eventually converge to a configuration in which none overlap. Naturally, there are limitations to such an initialization algorithm. For example, there exists thresholds for the solid area fraction (depending on the radius distribution) above which no circle packing is possible.
We have added a brief explanation of this initialization algorithm to the updated manuscript at the end of Section 3.2.
